---
title: "sgoli4-2 FML"
author: "Goli Sai Priyanka"
date: "2023-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## summary
Problem statement:

Universal bank is a young bank growing rapidly in terms of overall customer acquisition. The majority of these customers are liability customers (depositors) with varying sizes of relationship with the bank. The customer base of asset customers (borrowers) is quite small, and the bank is interested in expanding this base
rapidly in more loan business. In particular, it wants to explore ways of converting its liability customers to personal loan customers.
A campaign that the bank ran last year for liability customers showed a healthy conversion rate of over 9% success. This has encouraged the retail marketing department to devise smarter campaigns with better target marketing. The goal is to use k-NN to predict whether a new customer will accept a loan offer. This will serve as the basis for the design of a new campaign.
The file UniversalBank.csv contains data on 5000 customers. The data include customer demographic information (age, income, etc.), the customer’s relationship with the bank (mortgage, securities account, etc.), and the customer response to the last personal loan campaign (Personal Loan). Among these 5000
customers, only 480 (= 9.6%) accepted the personal loan that was offered to them in the earlier campaign. Partition the data into training (60%) and validation (40%) sets

```{r}
###Cleaning and Importing Data

 #Install and load the packages "class," "caret," and "e1071" first.

library(class)
library(caret)

```
```{r}
##Reading the data 
pp_data <- read.csv("C://Users//itspr//Downloads//UniversalBank.csv")
dim(pp_data)
```
```{r}
head(pp_data)
```
```{r}
tail(pp_data)
```
```{r}
t(t(names(pp_data))) ## We use the t function to transpose the dataframe.

```
```{r}
getwd()
```
```{r}
#Droping the  ID and ZIP
p.new_data <- pp_data[,-c(1,5)]
dim(p.new_data)
```
##40% should be used for validation and 60% for training. There are numerous approaches to accomplish this. Two approaches will be considered. Let us first convert category variables into dummy variables before splitting.

##altering the int value of the education attribute to char
```{r}
p.new_data$Education <- as.factor(p.new_data$Education)

```
#building the dummy variables for the "education" attribute.
```{r}
fake<- dummyVars(~.,data=p.new_data)
pp_data <- as.data.frame(predict(fake,p.new_data))
```

##We have to execute the function again, thus we need to create the seed and divide the data into training (60%) and validation (40%) groups.

```{r}
set.seed(1)
train.p_data <- sample(row.names(pp_data), 0.6*dim(pp_data)[1])
valids_data <- setdiff(row.names(pp_data),train.p_data)
train <- pp_data[train.p_data,]
valid <- pp_data[valids_data,]
t(t(names(train)))
```
```{r}
summary(train)
```
```{r}
cat("The size of the training dataset is:",nrow(train))
```
```{r}
summary(valid)
```

```{r}
cat("The size of the validation dataset is:",nrow(valid))
```

##lets normalize the data
```{r}
train.p_norm <- train[,-10]
valids_norm <- valid[,-10]
norm <- preProcess(train[,-10],method=c("center","scale"))
5
train.p_norm <- predict(norm,train[,-10])
valids_norm <- predict(norm,valid[,-10])
```
Questions
Consider the following customer:
1. Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities
Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform
a k-NN classification with all predictors except ID and ZIP code using k = 1.
Remember to transform categorical predictors with more than two categories
into dummy variables first. Specify the success class as 1 (loan acceptance), and
use the default cutoff value of 0.5. How would this customer be classified?


Creating new customer data:

```{r}
latest.customer <- data.frame(
Age = 40,
Experience = 10,
Income = 84,
Family = 2,
CCAvg = 2,
Education.1 = 0,
Education.2 = 1,
Education.3 = 0,
Mortgage = 0,
Securities.Account = 0,
CD.Account = 0,
Online = 1,
CreditCard = 1
)

# Normalize the new customer dataset
cust_latest.norm <- predict(norm, latest.customer)
```

We are Performing kNN classification
```{r}
ppredictions <- class::knn(train = train.p_norm,
test = cust_latest.norm,
cl = train$Personal.Loan, k = 1)
ppredictions
```

2.What is a choice of k that balances between overfitting and ignoring the predictor information?

```{r}
# Determine each value of k's accuracy.
#Determine the k-value range that will be taken into account.
accuracy <- data.frame(k = seq(1, 15, 1), overallaccuracy = rep(0, 15))
for(i in 1:15) {
kn <- class::knn(train = train.p_norm,
test = valids_norm,
cl = train$Personal.Loan, k = i)
accuracy[i, 2] <- confusionMatrix(kn,
as.factor(valid$Personal.Loan),positive = "1")$overall[1]
}
which(accuracy[,2] == max(accuracy[,2]))
```
```{r}
accuracy
```
The highest performing k value between 1 and 15 is 3.This k is the most accurate for 3, as it finds a middle ground between overfitting and ignoring predictions.

```{r}
plot(accuracy$k,accuracy$overallaccuracy)
```

3. Show the confusion matrix for the validation data that results from using the best k.
 

confusion matrix

```{r}
predics<- class::knn(train = train.p_norm,
test = valids_norm,
cl = train$Personal.Loan, k=3)
confusionMatrix(predics,as.factor(valid$Personal.Loan))
```

4. Consider the following customer: Age = 40, Experience = 10, Income =
84,Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3
= 0,Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and
CreditCard = 1. Classify the customer using the best k.


#Create a data frame with the same column names for customer user 2.

```{r}
latest.cust2 <- data.frame(
Age = 40,
Experience = 10,
Income = 84,
Family = 2,
CCAvg = 2,
Education.1 = 0,
Education.2 = 1,
Education.3 = 0,
Mortgage = 0,
Securities.Account = 0,
CD.Account = 0,
Online = 1,
CreditCard = 1)

#Normalizing the 2nd client dataset
latest.cust2_norm <- predict(norm,latest.cust2)
```

5. Repeating the process by partitioning the data into three parts -
50%, 30%, 20%,Apply the k-NN method with the k chosen above. Compare the
confusion matrix of the test set with that of the training and validation sets.
Comment on the differences and their reason.

```{r}
set.seed(500)
Train.p_Index <- sample(row.names(pp_data), .5*dim(pp_data)[1])#create train index
9
#create validation index
Valids.p_Index <- sample(setdiff(row.names(pp_data),Train.p_Index),.3*dim(pp_data)[1])
Test_Index =setdiff(row.names(pp_data),union(Train.p_Index,Valids.p_Index))#create test index
train.df <- pp_data[Train.p_Index,]
cat("The size of the new training dataset is:", nrow(train.df))

valid.df <- pp_data[Valids.p_Index, ]
cat("The size of the new validation dataset is:", nrow(valid.df))

test.df <- pp_data[Test_Index, ]
cat("The size of the new test dataset is:", nrow(test.df))
```
Normalizing the data

```{r}
norms.p_values <- preProcess(train.df[, -10], method=c("center", "scale"))
train.df.norm <- predict(norm, train.df[, -10])
valid.df.norm <- predict(norm, valid.df[, -10])
test.df.norm <- predict(norm ,test.df[,-10])
```

kNN processing and confusion matrix creation using training, testing, and validation data.
```{r}
predict.p_3 <- class::knn(train = train.df.norm,
test = test.df.norm,
cl = train.df$Personal.Loan, k=3)
confusionMatrix(predict.p_3,as.factor(test.df$Personal.Loan))
```

```{r}
predict.p_4 <- class::knn(train = train.df.norm,
test = valid.df.norm,
cl = train.df$Personal.Loan, k=3)
confusionMatrix(predict.p_4,as.factor(valid.df$Personal.Loan))
```

```{r}
predict.p_5 <- class::knn(train = train.df.norm,
test = train.df.norm,
cl = train.df$Personal.Loan, k=3)
confusionMatrix(predict.p_5,as.factor(train.df$Personal.Loan))
```

